---
title: "Refcard pandoc citeproc"
output: 
  github_document:
    toc: true
    number_sections: true
    html_preview: false
  html_document:
    toc: true
    number_sections: true
link-citations: true
bibliography: refs.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(knitr)
library(xml2)

knitr::opts_chunk$set(
  echo = TRUE
)

# Markdown pseudo-engine
knit_engines$set(
  markdown = function(options) {
    opt <- options
    opt$results <- "asis"
    engine_output(
      options = opt, 
      code = options$code,
      out = paste(options$code, collapse="\n")
      )
  }
)
```

This is a reminder (to me) how to use pandoc locator terms in (R)Markdown documents. Things like:

```{markdown}
Blah blah [@wf;@coleman].
```

or

```{markdown}
Blah blah [see @wf, pp. 33-35; also @coleman, chap. 1].
```

The terms themselves are locale-specific. They are defined in [CSL locale files](https://github.com/citation-style-language/locales) [@pandoc, chapter "Citation syntax"].



# General things

Worthwhile metadata options

- `link-citations: [true|false]` -- create hyperlinks leading to items in the reference list

Force reference list to appear somewhere else than the end of the document by placing

```
::: {#refs}
:::
```

# Locator terms

```{r terms, include=FALSE, eval=FALSE}
# Take term list from a Haskel source file

# Fetch the file if not present
fname <- "Locator.hs"
if(!file.exists(fname)) {
  download.file(
    "https://raw.githubusercontent.com/jgm/pandoc/master/src/Text/Pandoc/Citeproc/Locator.hs",
    dest = fname
  )
}

# Extract the terms
terms <- tail(readLines(fname), 50) |>
  paste(collapse="") |>
  trimws() |>
  (function(x) gsub(" +", "", x))() |>
  stringi::stri_extract_first(regex = "(?<=locatorTerms=\\[).*(?=\\])") |>
  (function(x) gsub('"', "", x))() |>
  (function(x) strsplit(x, ",")[[1]])()

# Fetch locale files if not present
loclist <- c("pl-PL", "en-US")
fnames <- paste0("locales-", loclist, ".xml")
for(i in seq(along=loclist)) {
  if(!file.exists(fnames[i]))
  download.file(
    paste0("https://raw.githubusercontent.com/citation-style-language/locales/master/", fnames[i]),
    destfile = fnames[i]
  )
}

# Extract the terms for the different locales
xml <- read_xml(fnames[1])
xml_find_all(xml, ".//d1:term[@name = 'book']/following::d1:term/preceding::d1:term[not(@form)]")
```



# Examples (English)

```{markdown}
see [@wf, book 1, chap. 1, pp. 33-35]
```

```{markdown}
see @wf [book 1]
```


```{markdown}
see @wf [chapter 1]
```

```{markdown}
see @wf [column 1]
```

```{markdown}
see @wf [figure 1]
```

```{markdown}
see @wf [folio 1]
```

```{markdown}
see @wf [number 1]
```

```{markdown}
see @wf [line 1]
```

```{markdown}
see @wf [note 1]
```

```{markdown}
see @wf [opus 1]
```

```{markdown}
see @wf [paragraph 1]
```

```{markdown}
see @wf [part 1]
```

```{markdown}
see @wf [section 1]
```

```{markdown}
see @wf [sub verbo 1]
```

```{markdown}
see @wf [verse 1]
```

```{markdown}
see @wf [volume 1]
```


## Protect locators with curly braces

```{markdown}
[@wf{ii, A, D-Z}, with suffix]
```

```{markdown}
[@wf, {chapter iv, vi-xi} with suffix here]
```


```{markdown}
@wf [chapter 33] or 
Wasserman says blah again [-@wf]
```






# Appendix A: BibTeX file {-}

```{r}
cat(readLines("refs.bib"), sep="\n")
```


# Appendix B: Software

```{r}
rmarkdown::pandoc_version()
system2("pandoc-citeproc", "--version", stdout=TRUE)
```



# Forced references 

The reference list is force-placed here instead of the end of the document:

::: {#refs}
:::


# References {-}